Resonant Scattering in a Uniform Sphere with Large Optical Depth

### 2021-12-20 ms.tex
Adding parentheses around equation numbers

\begin{abstract}
%TODO: 225 words! All done.
The radiation transfer equation for resonant scattering is solved in spherical geometry. A monochromatic source of photons is assumed at the center of a uniform sphere of constant gas density. The intensity within the sphere and at its surface is found in the limit of large line-center optical depth using the Eddington approximation for the angular dependence. Two limits are studied: a steady source and an impulsive source. For the steady-state problem, the solution can be represented as a sum of two terms: a previously-known analytic solution of the equation with  mean intensity $J=0$ at the surface, and a novel, semi-analytic solution which enforces the correct boundary condition of zero-ingoing intensity at the surface. This solution is compared to that of the Monte-Carlo method, which is valid at arbitrary optical depth. It is shown that the size of the correction is order unity when the spectral peaks approach the Doppler core and decreases slowly with line center optical depth, which may explain discrepancies in previous studies. For the impulsive problem, the time, spatial, and frequency dependence of the solution are expressed using an eigenfunction expansion in order to characterize the escape time distribution and emergent spectra of photons. It is shown that the lowest-order eigenfrequency agrees well with the decay rate found in the Monte Carlo escape time distribution at sufficiently large line-center optical depths.

@@

Appendix \ref{app:rteqn_derivation} contains a derivation of the transfer equation for convenience. Starting with the full transfer equation, Equation \ref{eq:finaleqn}, ignoring photon destruction and including a photon emission term given by Equation \ref{eq:jem}, the steady-state transfer equation is
\be
\nabla^2 J + \left( \frac{k}{\Delta} \right)^2 \frac{\partial^2 J}{\partial \sigma^2} & = & 
- \frac{ \sqrt{6} kL}{4\pi \Delta^2} \delta^3(\vec{x} - \vec{x}_s) \delta (\sigma - \sigma_{\rm s}).
\label{eq:rt_no_destr}
\ee
where $J$ is the mean intensity, the spatial variable is $\vec{x}$, and $\vec{x}_s$ is the position of the source. We will consider only the case where $\vec{x}_s=0$. Following \citet{1973MNRAS.162...43H}, we have used a change of variables from $x$ to $\sigma$,
\be \label{eq:int_change_of_variables}
\sigma(x) = \sqrt{\frac{2}{3}}\int_0^x \frac{dx}{\phi(x) \Delta} \approx \sqrt{\frac{2}{3}}\frac{\pi}{a}\frac{x^3}{3}, 
\ee
where the approximation is applicable on the damping wing. The line profile is then approximately (Equation \ref{eq:app:line_profile_wing})
\be \label{eq:line_profile_approx}
\phi \approx \frac{a}{\pi x^2 \Delta} \approx \frac{1}{3 \Delta}\left(\frac{2a}{\pi}\right)^{1/3}|\sigma|^{-2/3}.
\ee
In Equation \ref{eq:rt_no_destr}, $\sigma_s \equiv \sigma(x_s)$ is the photon frequency of the source. $\sigma_s$ is interchangeable with $x_s$ and $\nu_s$ in Doppler widths or Hz, respectively. Balancing the two terms on the left hand side of Equation \ref{eq:rt_no_destr} gives $\sigma \ {\sim}\ \tau_0$, or $x_{\rm peak}\ {\sim}\ (a\tau_0)^{1/3}$. The boundary condition of no incoming intensity at the surface \citep{1986rpa..book.....R} is
\be
J & = & \sqrt{3} H
\label{eq:bc}
\ee
at $r=R$.

@@

A better approximation to the true solution has been derived by \citet{2006ApJ...649...14D}, who generalized the closed-form solution in slab geometry found in \citet{1990ApJ...350..216N}. It satisfies a $J=0$ boundary condition at $r=R$. Again, we generalize their solution to allow emission at frequency $\sigma_{\rm s}$ away from line center. The result can be written as a sum over spatial modes,
\be \label{eq:J0_sum}
J_0 = \frac{\sqrt{6}L}{16\pi \Delta} \frac{1}{R^2}\sum_{n=1}^{\infty}n\frac{\sin{\kappa_n r}}{\kappa_n r}\exp{\left(\frac{-\kappa_n \Delta}{k}|\sigma - \sigma_s|\right)},
\ee
and
\be \label{eq:H0_sum}
H_0 = - \frac{1}{3k\phi} \frac{\partial J_0}{\partial r} = -\frac{1}{3k\phi}\frac{\sqrt{6}L}{16\pi\Delta} \frac{1}{R^2}\sum_{n=1}^{\infty}n\left(\frac{\cos{\kappa_n r}}{r} - \frac{\sin{\kappa_n r}}{\kappa_n r^2}\right)\exp{\left(\frac{-\kappa_n \Delta}{k}|\sigma - \sigma_s|\right)},
\ee
where $\kappa_n=n\pi/R$. These can be summed to give the closed form expressions
\be
J_0 & = & \frac{\sqrt{6}L}{32\pi^2 \Delta}
\frac{1}{Rr}
\left( 
\frac{ \sin(\pi r/R) }{ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] - \cos(\pi r/R)}
\right)
\label{eq:J0}
\ee
and
\be
H_0 & = &\frac{1}{3k\phi}
\frac{\sqrt{6}L}{32\pi^2 \Delta}
\frac{1}{Rr^2}
\left( 
\frac{ \sin(\pi r/R) }{ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] - \cos(\pi r/R)}
\right. \nonumber \\ & & \left. - \left( \frac{\pi r}{R} \right)
\frac{ \cos(\pi r/R) }{ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] - \cos(\pi r/R)}
+ \left( \frac{\pi r}{R} \right)
\frac{ \sin^2(\pi r/R) }{ \left[ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] - \cos(\pi r/R) \right]^2 }
\right).
\label{eq:H0}
\ee
These solutions agree with Equations \ref{eq:Jd} and \ref{eq:Hd} when the arguments of the trigonometric and hyperbolic functions are small. Again $J_0 \gg H_0$, except near $r=R$, where it goes to zero. The flux at $r=R$ can be written
\be
\nonumber
H_0(R, \sigma) & = & - \frac{1}{3k\phi}
\frac{\sqrt{6}L}{16\pi \Delta}
\frac{1}{R^3}
\sum_{n=1}^{\infty} 
n (-1)^n \exp{\left(\frac{-\kappa_n \Delta}{k}|\sigma - \sigma_s|\right)}\\
& = &  \frac{1}{3k\phi}
\frac{\sqrt{6}L}{32\pi \Delta}
\frac{1}{R^3}
\left( 
\frac{ 1 }{ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] +1 }
\right).
\label{eq:H0surf}
\ee
Equation \ref{eq:H0surf} will be shown to be a better approximation to the solution than Equation \ref{eq:Hd}. It is still valid near the delta function at $r=0$, but is also a better approximation at $r=R$. $J_0$ decreases exponentially rather than as a power-law in frequency as it does for $J_d$, giving a much smaller flux on the line wings as compared to the divergent solution.

In order to enforce the boundary conditions, a different solution method is attempted here, namely a continuous Fourier expansion in the frequency variable $\sigma$. The solution of this problem is split into two pieces: $J_0$ which includes the delta function source and satisfies $J=0$ at $r=R$, and $J_{\rm bc}$ which allows the boundary condition $J=\sqrt{3}H$ to be satisfied at $r=R$. The total solution is
\be
J(r,\sigma) & = & J_0(r,\sigma) + J_{\rm bc}(r,\sigma)
\ee
and
\be \label{eq:totalflux}
H(r,\sigma) & = & H_0(r,\sigma) + H_{\rm bc}(r,\sigma).
\ee
The additional term $J_{\rm bc}$ must then be a solution of the homogeneous equation
\be \label{eq:diffeq}
\frac{\partial^2J_{\rm bc}}{\partial r^2} + \frac{2}{r} \frac{\partial J_{\rm bc}}{\partial r}
+ \left( \frac{k}{\Delta} \right)^2 \frac{\partial^2 J_{\rm bc}}{\partial \sigma^2} &= & 0
\ee
with no delta function source term, and it must allow the boundary conditions to be satisfied at the surface. Since $J_0(R,
\sigma)=0$, the surface boundary condition becomes
\be
J_{\rm bc}(R,\sigma) - \sqrt{3} H_{\rm bc}(R,\sigma) & = 
\sqrt{3} H_0(R,\sigma).
\label{eq:bc2}
\ee
Inserting a frequency dependence $J_{\rm bc} \propto e^{is\sigma}$, for ``wavenumber" $s$, gives the equation for modified spherical Bessel functions of the first kind, $i_0(z)=\sinh(z)/z$ for the radial dependence. The solution can then be represented as
\be
J_{\rm bc}(r,\sigma) & = & 
\int_{-\infty}^\infty \frac{ds}{2\pi} e^{is\sigma} A(s) 
\frac{i_0(krs/\Delta)}{i_0(kRs/\Delta)},
\label{eq:Jbc}
\ee
where $A(s)$ is the Fourier amplitude. Inserting Equation \ref{eq:Jbc} into Equation \ref{eq:bc2} leads to the following equation for the Fourier amplitudes,
\be
\int_{-\infty}^\infty \frac{ds}{2\pi} e^{is\sigma} A(s)
\left[ 1 + \left( \frac{s}{\sqrt{3} \Delta \phi} \right) \left( \frac{i_0^\prime(kRs/\Delta)}{i_0(kRs/\Delta)} \right) \right]
& = & \sqrt{3} H_0(R,\sigma).
\label{eq:bc3}
\ee
Discretization of Equation \ref{eq:bc3} for frequency variables $\sigma_i$ and wavenumbers $s_j$
leads to a set of coupled linear equations for the $A(s_j)$. We use equally-spaced points $\delta \sigma = 2\sigma_{\rm max}/(N-1)$ and $\delta s = 2\pi/(N\delta \sigma)$, where $N$ is the number of points for each grid. The result is a linear system to solve for the $A(s)$. The maximum frequency is set as $\sigma_{\rm max} = {\rm constant} \times \tau_0$, for a large enough constant that the end of the frequency grid is at such small intensities that it does not affect the solution except close to the boundaries. The number of points was increased until the solution was well-resolved near line center, and only became inaccurate close to the boundaries. Given the Fourier amplitudes $A(s)$, $J_{\rm bc}$ is computed using Equation \ref{eq:Jbc}, and the flux is given by
\be
H_{\rm bc}(r,\sigma) & = & -\frac{1}{3k\phi}
\frac{\partial J_{\rm bc}(r,\sigma)}{\partial r}
= -\frac{1}{3k\phi}
\int_{-\infty}^\infty \frac{ds}{2\pi} e^{is\sigma} A(s) 
\left( \frac{ks}{\Delta} \right) 
\left( \frac{i_0^\prime(krs/\Delta)}{i_0(kRs/\Delta)} \right).
\label{eq:Hbc}
\ee
The Bessel functions are finite at the center and rise steeply toward the surface when $kRs/\Delta \gg 1$.

@@

We now estimate the scaling of $H_{\rm bc}$ with $\tau_0$. In the limit $J_{\rm bc} \gg H_{\rm bc}$, we find that $J_{\rm bc} \approx \sqrt{3} H_0$ from Equation \ref{eq:bc2}. We  calculate $H_{\rm bc}$ from $J_{\rm bc}$ using Equation \ref{eq:Hbc} as
\be
H_{\rm bc}(R, \sigma) \approx \frac{1}{\sqrt{3}k\phi}\frac{ks}{\Delta}H_0\ {\sim}\ H_0 s \frac{x^2}{a}\ {\sim}\ H_0 \frac{1}{\tau_0}\frac{(a\tau_0)^{2/3}}{a}\ {\sim}\ H_0 (a\tau_0)^{-1/3},
\ee
where we have used $s\ {\sim}\ 1/\sigma\ {\sim}\ 1/\tau_0$ so that
\be \label{eq:hbc_scaling}
\frac{H_{\rm bc}(R, \sigma)}{H_0(R, \sigma)} \propto (a\tau_0)^{-1/3}.
\ee
At large $\tau_0$, it is expected that the correction term will be small, but it will become increasingly important as $\tau_0$ decreases. Our solution of the transfer equation is only valid when the peaks of the spectral energy distribution lie well outside of the Doppler core, e.g., for large $\tau_0$. Thus, the value of $x$ at which the Doppler and Lorentzian components of the line profile are equal is $x_{\rm cw}=3.3$. Setting $x_{\rm cw} = x_{\rm peak}$ and solving for $\tau_0$ gives the value at which the peak of the spectrum falls at the Doppler core boundary, which is $\tau_{\rm cp} \approx 10^5$ (``core-peak'' optical depth). Hence $H_{\rm bc}/H_0 {\sim} (\tau_{\rm cp}/\tau_0)^{1/3}$ is large at $\tau_0 \leq \tau_{\rm cp}$ and decreases relatively slowly as $\tau_0$ increases. Additionally, the optical depth at $x_{\rm peak}$ is proportional to $(a\tau_0)^{1/3}$, so photons here become optically thin when $a\tau_0 {\sim} 1$.

@@

In Figure \ref{fig:sol_mc_tau} we show the solutions alongside Monte Carlo for $\tau_0=10^5, 10^6$, and $10^7$. From Equation \ref{eq:hbc_scaling}, the size of the term $H_{\rm bc}$ should become smaller with larger optical depths, following a $(a\tau_0)^{-1/3}$ scaling. Indeed, agreement between Equation \ref{eq:H0surf} and the Monte Carlo points in Fig. \ref{fig:sol_mc_tau} improves as $\tau_0$ increases, with $H_{\rm bc}$ providing a fractionally smaller correction to $H_0$. One factor of $(a\tau_0)^{1/3}$ has been scaled out of the x-axis such that the peaks of the distributions are horizontally aligned. This scaling has also been applied to the y-axis to preserve normalization of the escape probability. At lower $\tau_0$, the scattering of photons within the Doppler core of the line becomes important, but our analytic solution does not include this effect. The effects of line core scattering can be seen in the Monte Carlo data for $\tau_0=10^5$ and, to a lesser extent, $
\tau_0=10^6$.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{xinit.pdf}
    \caption{$P_{\rm bc}(x) = 16\pi^2R^2H_{\rm bc}(R, x)\Delta/L$ (Equation \ref{eq:prob_spectrum} with $H \rightarrow H_{\rm bc}$) vs. frequency $x$. Thicker lines indicate larger optical depth labeled by $\tau_0=10^5, 10^7,$ and $10^9$, and darker lines indicate $\sigma_s$ further from line center, labeled by $\sigma_s=0.0, \tau_0, 2\tau_0$. The location of the core-wing boundary $x_{cw}$ is shown with green vertical lines of increasing thickness for each $\tau_0$. The axes are scaled to show that the size of the correction factor agrees with the predicted $(a\tau_0)^{-1/3}$ scaling.}
    \label{fig:xinit}
\end{figure}

Figure \ref{fig:xinit} shows how the correction $P_{\rm bc}(x)$ (Equation \ref{eq:prob_spectrum} with $H \rightarrow H_{\rm bc}$) scales with both $\tau_0$ and $x_s$. In this figure, $\sigma_s$ is shifted by integer multiples of $\tau_0$ (Equation \ref{eq:change_of_variables}) such that the source falls near the peak of the spectrum for each $\tau_0$. For clarity, only the $x > 0$ side of the spectrum is shown. From Equation \ref{eq:hbc_scaling}, it is expected that the fractional size of $H_{\rm bc}$ relative to $H_0$ should become smaller with larger optical depths following $(a\tau_0)^{-1/3}$. This factor has been scaled out of the figure such that solutions for different $\tau_0$ and the same $\sigma_s$ should show close agreement in scale on the figure's vertical axis if the relation holds. Indeed, the scaled solutions converge as $\tau_0$ becomes larger, indicating agreement with the $(a\tau_0)^{-1/3}$ scaling. The remaining discrepancy present in the vertical axis for fixed $\sigma_s$ results from $x_{\rm peak}$ becoming close to $x_{\rm cw}$; at lower $\tau_0$, this causes the line profile approximation in the wing (Equation \ref{eq:app:line_profile_wing}) to break down. From this, we conclude that the errors introduced by the incorrect separation of variables in \citet{1973MNRAS.162...43H}, \citet{1990ApJ...350..216N}, \citet{2006ApJ...649...14D} and others are indeed proportional to $(a\tau_0)^{-1/3}$.

@@

We now calculate the distribution of the times it takes each photon to escape the sphere. This is obtained by integrating Equation \ref{eq:dEdtdnu} over all frequencies. We find
\be
P(t)  & = & \sqrt{\frac{3}{2}} \frac{16\pi^2 R \Delta^2 }{3kE}     \sum_{nm} (-1)^{n+1}  e^{-\gamma_{nm}t} \int d\sigma J_{nm}(\sigma) 
\nonumber \\ & = &  \sum_{nm} P_{nm} \gamma_{nm} e^{-\gamma_{nm}t},
\label{eq:waittime}
\ee
which is normalized to unity. For a sufficiently large number of spatial modes $n$ and frequency modes $m$, the result of this sum can agree with Monte Carlo escape time distributions when $x_s=0$. The late-time distribution is simply an exponential falloff. The rate constant of the falloff is the lowest-order eigenfrequency, $\gamma_{11}$, and its scale is determined by the coefficient $P_{11}$ as in Equation \ref{eq:pnmsoln}. Thus, an approximate ``fitting function'' that captures both the peak of the escape time distribution and the exponential falloff is
\be \label{eq:fitting_function}
P(t) = \exp{\left[-\left(\frac{t_{\rm diff}}{t}\right)^2\right]} \times \gamma_{11} P_{11} e^{-\gamma_{11}t}.
\ee
The first term represents the early-time distribution, which then transitions to an exponential falloff past a point $ct_{\rm diff}/R = (a\tau_0)^{1/3}$, where $t_{\rm diff}$ is the characteristic diffusion timescale.

>>>>

\begin{abstract}
The radiation transfer equation for resonant scattering is solved in spherical geometry. A monochromatic source of photons is assumed at the center of a uniform sphere of constant gas density. The intensity within the sphere and at its surface is found in the limit of large line-center optical depth using the Eddington approximation for the angular dependence. Two limits are studied: a steady source and an impulsive source. For the steady-state problem, the solution can be represented as a sum of two terms: a previously-known analytic solution of the equation with  mean intensity $J=0$ at the surface, and a novel, semi-analytic solution which enforces the correct boundary condition of zero-ingoing intensity at the surface. This solution is compared to that of the Monte-Carlo method, which is valid at arbitrary optical depth. It is shown that the size of the correction is order unity when the spectral peaks approach the Doppler core and decreases slowly with line center optical depth, which may explain discrepancies in previous studies. For the impulsive problem, the time, spatial, and frequency dependence of the solution are expressed using an eigenfunction expansion in order to characterize the escape time distribution and emergent spectra of photons. It is shown that the lowest-order eigenfrequency agrees well with the decay rate found in the Monte Carlo escape time distribution at sufficiently large line-center optical depths.

@@

Appendix \ref{app:rteqn_derivation} contains a derivation of the transfer equation for convenience. Starting with the full transfer equation, Equation (\ref{eq:finaleqn}), ignoring photon destruction and including a photon emission term given by Equation (\ref{eq:jem}), the steady-state transfer equation is
\be
\nabla^2 J + \left( \frac{k}{\Delta} \right)^2 \frac{\partial^2 J}{\partial \sigma^2} & = & 
- \frac{ \sqrt{6} kL}{4\pi \Delta^2} \delta^3(\vec{x} - \vec{x}_s) \delta (\sigma - \sigma_{\rm s}).
\label{eq:rt_no_destr}
\ee
where $J$ is the mean intensity, the spatial variable is $\vec{x}$, and $\vec{x}_s$ is the position of the source. We will consider only the case where $\vec{x}_s=0$. Following \citet{1973MNRAS.162...43H}, we have used a change of variables from $x$ to $\sigma$,
\be \label{eq:int_change_of_variables}
\sigma(x) = \sqrt{\frac{2}{3}}\int_0^x \frac{dx}{\phi(x) \Delta} \approx \sqrt{\frac{2}{3}}\frac{\pi}{a}\frac{x^3}{3}, 
\ee
where the approximation is applicable on the damping wing. From Equation (\ref{eq:app:line_profile_wing}), the line profile is then approximately 
\be \label{eq:line_profile_approx}
\phi \approx \frac{a}{\pi x^2 \Delta} \approx \frac{1}{3 \Delta}\left(\frac{2a}{\pi}\right)^{1/3}|\sigma|^{-2/3}.
\ee
In Equation (\ref{eq:rt_no_destr}), $\sigma_s \equiv \sigma(x_s)$ is the photon frequency of the source. $\sigma_s$ is interchangeable with $x_s$ and $\nu_s$ in Doppler widths or Hz, respectively. Balancing the two terms on the left hand side of Equation (\ref{eq:rt_no_destr}) gives $\sigma \ {\sim}\ \tau_0$, or $x_{\rm peak}\ {\sim}\ (a\tau_0)^{1/3}$. The boundary condition of no incoming intensity at the surface \citep{1986rpa..book.....R} is
\be
J & = & \sqrt{3} H
\label{eq:bc}
\ee
at $r=R$.

@@

A better approximation to the true solution has been derived by \citet{2006ApJ...649...14D}, who generalized the closed-form solution in slab geometry found in \citet{1990ApJ...350..216N}. It satisfies a $J=0$ boundary condition at $r=R$. Again, we generalize their solution to allow emission at frequency $\sigma_{\rm s}$ away from line center. The result can be written as a sum over spatial modes,
\be \label{eq:J0_sum}
J_0 = \frac{\sqrt{6}L}{16\pi \Delta} \frac{1}{R^2}\sum_{n=1}^{\infty}n\frac{\sin{\kappa_n r}}{\kappa_n r}\exp{\left(\frac{-\kappa_n \Delta}{k}|\sigma - \sigma_s|\right)},
\ee
and
\be \label{eq:H0_sum}
H_0 = - \frac{1}{3k\phi} \frac{\partial J_0}{\partial r} = -\frac{1}{3k\phi}\frac{\sqrt{6}L}{16\pi\Delta} \frac{1}{R^2}\sum_{n=1}^{\infty}n\left(\frac{\cos{\kappa_n r}}{r} - \frac{\sin{\kappa_n r}}{\kappa_n r^2}\right)\exp{\left(\frac{-\kappa_n \Delta}{k}|\sigma - \sigma_s|\right)},
\ee
where $\kappa_n=n\pi/R$. These can be summed to give the closed form expressions
\be
J_0 & = & \frac{\sqrt{6}L}{32\pi^2 \Delta}
\frac{1}{Rr}
\left( 
\frac{ \sin(\pi r/R) }{ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] - \cos(\pi r/R)}
\right)
\label{eq:J0}
\ee
and
\be
H_0 & = &\frac{1}{3k\phi}
\frac{\sqrt{6}L}{32\pi^2 \Delta}
\frac{1}{Rr^2}
\left( 
\frac{ \sin(\pi r/R) }{ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] - \cos(\pi r/R)}
\right. \nonumber \\ & & \left. - \left( \frac{\pi r}{R} \right)
\frac{ \cos(\pi r/R) }{ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] - \cos(\pi r/R)}
+ \left( \frac{\pi r}{R} \right)
\frac{ \sin^2(\pi r/R) }{ \left[ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] - \cos(\pi r/R) \right]^2 }
\right).
\label{eq:H0}
\ee
These solutions agree with Equations (\ref{eq:Jd}) and (\ref{eq:Hd}) when the arguments of the trigonometric and hyperbolic functions are small. Again $J_0 \gg H_0$, except near $r=R$, where it goes to zero. The flux at $r=R$ can be written
\be
\nonumber
H_0(R, \sigma) & = & - \frac{1}{3k\phi}
\frac{\sqrt{6}L}{16\pi \Delta}
\frac{1}{R^3}
\sum_{n=1}^{\infty} 
n (-1)^n \exp{\left(\frac{-\kappa_n \Delta}{k}|\sigma - \sigma_s|\right)}\\
& = &  \frac{1}{3k\phi}
\frac{\sqrt{6}L}{32\pi \Delta}
\frac{1}{R^3}
\left( 
\frac{ 1 }{ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] +1 }
\right).
\label{eq:H0surf}
\ee
Equation (\ref{eq:H0surf}) will be shown to be a better approximation to the solution than Equation (\ref{eq:Hd}). It is still valid near the delta function at $r=0$, but is also a better approximation at $r=R$. $J_0$ decreases exponentially rather than as a power-law in frequency as it does for $J_d$, giving a much smaller flux on the line wings as compared to the divergent solution.

In order to enforce the boundary conditions, a different solution method is attempted here, namely a continuous Fourier expansion in the frequency variable $\sigma$. The solution of this problem is split into two pieces: $J_0$ which includes the delta function source and satisfies $J=0$ at $r=R$, and $J_{\rm bc}$ which allows the boundary condition $J=\sqrt{3}H$ to be satisfied at $r=R$. The total solution is
\be
J(r,\sigma) & = & J_0(r,\sigma) + J_{\rm bc}(r,\sigma)
\ee
and
\be \label{eq:totalflux}
H(r,\sigma) & = & H_0(r,\sigma) + H_{\rm bc}(r,\sigma).
\ee
The additional term $J_{\rm bc}$ must then be a solution of the homogeneous equation
\be \label{eq:diffeq}
\frac{\partial^2J_{\rm bc}}{\partial r^2} + \frac{2}{r} \frac{\partial J_{\rm bc}}{\partial r}
+ \left( \frac{k}{\Delta} \right)^2 \frac{\partial^2 J_{\rm bc}}{\partial \sigma^2} &= & 0
\ee
with no delta function source term, and it must allow the boundary conditions to be satisfied at the surface. Since $J_0(R,
\sigma)=0$, the surface boundary condition becomes
\be
J_{\rm bc}(R,\sigma) - \sqrt{3} H_{\rm bc}(R,\sigma) & = 
\sqrt{3} H_0(R,\sigma).
\label{eq:bc2}
\ee
Inserting a frequency dependence $J_{\rm bc} \propto e^{is\sigma}$, for ``wavenumber" $s$, gives the equation for modified spherical Bessel functions of the first kind, $i_0(z)=\sinh(z)/z$ for the radial dependence. The solution can then be represented as
\be
J_{\rm bc}(r,\sigma) & = & 
\int_{-\infty}^\infty \frac{ds}{2\pi} e^{is\sigma} A(s) 
\frac{i_0(krs/\Delta)}{i_0(kRs/\Delta)},
\label{eq:Jbc}
\ee
where $A(s)$ is the Fourier amplitude. Inserting Equation (\ref{eq:Jbc}) into Equation (\ref{eq:bc2}) leads to the following equation for the Fourier amplitudes,
\be
\int_{-\infty}^\infty \frac{ds}{2\pi} e^{is\sigma} A(s)
\left[ 1 + \left( \frac{s}{\sqrt{3} \Delta \phi} \right) \left( \frac{i_0^\prime(kRs/\Delta)}{i_0(kRs/\Delta)} \right) \right]
& = & \sqrt{3} H_0(R,\sigma).
\label{eq:bc3}
\ee
Discretization of Equation (\ref{eq:bc3}) for frequency variables $\sigma_i$ and wavenumbers $s_j$
leads to a set of coupled linear equations for the $A(s_j)$. We use equally-spaced points $\delta \sigma = 2\sigma_{\rm max}/(N-1)$ and $\delta s = 2\pi/(N\delta \sigma)$, where $N$ is the number of points for each grid. The result is a linear system to solve for the $A(s)$. The maximum frequency is set as $\sigma_{\rm max} = {\rm constant} \times \tau_0$, for a large enough constant that the end of the frequency grid is at such small intensities that it does not affect the solution except close to the boundaries. The number of points was increased until the solution was well-resolved near line center, and only became inaccurate close to the boundaries. Given the Fourier amplitudes $A(s)$, $J_{\rm bc}$ is computed using Equation (\ref{eq:Jbc}), and the flux is given by
\be
H_{\rm bc}(r,\sigma) & = & -\frac{1}{3k\phi}
\frac{\partial J_{\rm bc}(r,\sigma)}{\partial r}
= -\frac{1}{3k\phi}
\int_{-\infty}^\infty \frac{ds}{2\pi} e^{is\sigma} A(s) 
\left( \frac{ks}{\Delta} \right) 
\left( \frac{i_0^\prime(krs/\Delta)}{i_0(kRs/\Delta)} \right).
\label{eq:Hbc}
\ee
The Bessel functions are finite at the center and rise steeply toward the surface when $kRs/\Delta \gg 1$.

@@

We now estimate the scaling of $H_{\rm bc}$ with $\tau_0$. In the limit $J_{\rm bc} \gg H_{\rm bc}$, we find that $J_{\rm bc} \approx \sqrt{3} H_0$ from Equation (\ref{eq:bc2}). We  calculate $H_{\rm bc}$ from $J_{\rm bc}$ using Equation (\ref{eq:Hbc}) as
\be
H_{\rm bc}(R, \sigma) \approx \frac{1}{\sqrt{3}k\phi}\frac{ks}{\Delta}H_0\ {\sim}\ H_0 s \frac{x^2}{a}\ {\sim}\ H_0 \frac{1}{\tau_0}\frac{(a\tau_0)^{2/3}}{a}\ {\sim}\ H_0 (a\tau_0)^{-1/3},
\ee
where we have used $s\ {\sim}\ 1/\sigma\ {\sim}\ 1/\tau_0$ so that
\be \label{eq:hbc_scaling}
\frac{H_{\rm bc}(R, \sigma)}{H_0(R, \sigma)} \propto (a\tau_0)^{-1/3}.
\ee
At large $\tau_0$, it is expected that the correction term will be small, but it will become increasingly important as $\tau_0$ decreases. Our solution of the transfer equation is only valid when the peaks of the spectral energy distribution lie well outside of the Doppler core, e.g., for large $\tau_0$. Thus, the value of $x$ at which the Doppler and Lorentzian components of the line profile are equal is $x_{\rm cw}=3.3$. Setting $x_{\rm cw} = x_{\rm peak}$ and solving for $\tau_0$ gives the value at which the peak of the spectrum falls at the Doppler core boundary, which is $\tau_{\rm cp} \approx 10^5$ (``core-peak'' optical depth). Hence $H_{\rm bc}/H_0 {\sim} (\tau_{\rm cp}/\tau_0)^{1/3}$ is large at $\tau_0 \leq \tau_{\rm cp}$ and decreases relatively slowly as $\tau_0$ increases. Additionally, the optical depth at $x_{\rm peak}$ is proportional to $(a\tau_0)^{1/3}$, so photons here become optically thin when $a\tau_0 {\sim} 1$.

@@

In Figure \ref{fig:sol_mc_tau} we show the solutions alongside Monte Carlo for $\tau_0=10^5, 10^6$, and $10^7$. From Equation (\ref{eq:hbc_scaling}), the size of the term $H_{\rm bc}$ should become smaller with larger optical depths, following a $(a\tau_0)^{-1/3}$ scaling. Indeed, agreement between Equation (\ref{eq:H0surf}) and the Monte Carlo points in Fig. \ref{fig:sol_mc_tau} improves as $\tau_0$ increases, with $H_{\rm bc}$ providing a fractionally smaller correction to $H_0$. One factor of $(a\tau_0)^{1/3}$ has been scaled out of the x-axis such that the peaks of the distributions are horizontally aligned. This scaling has also been applied to the y-axis to preserve normalization of the escape probability. At lower $\tau_0$, the scattering of photons within the Doppler core of the line becomes important, but our analytic solution does not include this effect. The effects of line core scattering can be seen in the Monte Carlo data for $\tau_0=10^5$ and, to a lesser extent, $
\tau_0=10^6$.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{xinit.pdf}
    \caption{$P_{\rm bc}(x) = 16\pi^2R^2H_{\rm bc}(R, x)\Delta/L$ (Equation (\ref{eq:prob_spectrum}) with $H \rightarrow H_{\rm bc}$) vs. frequency $x$. Thicker lines indicate larger optical depth labeled by $\tau_0=10^5, 10^7,$ and $10^9$, and darker lines indicate $\sigma_s$ further from line center, labeled by $\sigma_s=0.0, \tau_0, 2\tau_0$. The location of the core-wing boundary $x_{cw}$ is shown with green vertical lines of increasing thickness for each $\tau_0$. The axes are scaled to show that the size of the correction factor agrees with the predicted $(a\tau_0)^{-1/3}$ scaling.}
    \label{fig:xinit}
\end{figure}

Figure \ref{fig:xinit} shows how the correction $P_{\rm bc}(x)$, Equation (\ref{eq:prob_spectrum}) with $H \rightarrow H_{\rm bc}$, scales with both $\tau_0$ and $x_s$. In this figure, $\sigma_s$ is shifted by integer multiples of $\tau_0$ (Equation \ref{eq:change_of_variables}) such that the source falls near the peak of the spectrum for each $\tau_0$. For clarity, only the $x > 0$ side of the spectrum is shown. From Equation (\ref{eq:hbc_scaling}), it is expected that the fractional size of $H_{\rm bc}$ relative to $H_0$ should become smaller with larger optical depths following $(a\tau_0)^{-1/3}$. This factor has been scaled out of the figure such that solutions for different $\tau_0$ and the same $\sigma_s$ should show close agreement in scale on the figure's vertical axis if the relation holds. Indeed, the scaled solutions converge as $\tau_0$ becomes larger, indicating agreement with the $(a\tau_0)^{-1/3}$ scaling. The remaining discrepancy present in the vertical axis for fixed $\sigma_s$ results from $x_{\rm peak}$ becoming close to $x_{\rm cw}$; at lower $\tau_0$, this causes the line profile approximation in the wing, Equation (\ref{eq:app:line_profile_wing}), to break down. From this, we conclude that the errors introduced by the incorrect separation of variables in \citet{1973MNRAS.162...43H}, \citet{1990ApJ...350..216N}, \citet{2006ApJ...649...14D} and others are indeed proportional to $(a\tau_0)^{-1/3}$.

@@

We now calculate the distribution of the times it takes each photon to escape the sphere. This is obtained by integrating Equation (\ref{eq:dEdtdnu}) over all frequencies. We find
\be
P(t)  & = & \sqrt{\frac{3}{2}} \frac{16\pi^2 R \Delta^2 }{3kE}     \sum_{nm} (-1)^{n+1}  e^{-\gamma_{nm}t} \int d\sigma J_{nm}(\sigma) 
\nonumber \\ & = &  \sum_{nm} P_{nm} \gamma_{nm} e^{-\gamma_{nm}t},
\label{eq:waittime}
\ee
which is normalized to unity. For a sufficiently large number of spatial modes $n$ and frequency modes $m$, the result of this sum can agree with Monte Carlo escape time distributions when $x_s=0$. The late-time distribution is simply an exponential falloff. The rate constant of the falloff is the lowest-order eigenfrequency, $\gamma_{11}$, and its scale is determined by the coefficient $P_{11}$ as in Equation (\ref{eq:pnmsoln}). Thus, an approximate ``fitting function'' that captures both the peak of the escape time distribution and the exponential falloff is
\be \label{eq:fitting_function}
P(t) = \exp{\left[-\left(\frac{t_{\rm diff}}{t}\right)^2\right]} \times \gamma_{11} P_{11} e^{-\gamma_{11}t}.
\ee
The first term represents the early-time distribution, which then transitions to an exponential falloff past a point $ct_{\rm diff}/R = (a\tau_0)^{1/3}$, where $t_{\rm diff}$ is the characteristic diffusion timescale.
